Cryptanalysis of SBC based post-quantum signature scheme \\\textit{Graduation internship report}

### 2025-07-23 rapport.tex
correction de Sorina à faire

	\newpage
	\tableofcontents
	\newpage
	\section{MPCitH and SBC signature scheme}\label{sec2}

@@

		The \emph{MPC-in-the-Head} (MPCitH) framework, introduced by \cite{IKOS07}, is a technique for constructing zero-knowledge proofs. It enables a prover to convince a verifier that they possess a secret witness $w$ for a given public input $x$, without revealing any information about the witness itself.

		Given $\mathcal{L}$ a language in NP (i.e. $x \in \mathcal{L} \Longleftrightarrow \exists w, \mathcal{R}(x, w) = 1$). Given a public $x$, $\mathcal{P}$ wants to prove that he knows $w$ such that: $\mathcal{R}(x, w)$.

@@

		To represent the secret witness $w$, a linear secret sharing scheme is used among $N$ simulated parties. Each party $i$ receives a random share denoted by $[w]^{\lir}$, and the prover computes an additional correction term defined as:
		$$
		\delta_w = w - \sum_{i=1}^{N} [w]^{\lir}.
		$$
		This ensures that the full witness can be correctly reconstructed from the shares and the correction term during the simulated execution.

@@

		We introduce some standard notations in MPCitH protocols. 
		\begin{itemize}
			\item $[w]$, $w$ is a shared value.
			\item $w^{\lir}$ is the $i$-th share of w.
			\item $w^{[e]}$ is a value used in the $e$-th round of the protocol.
			\item $\delta_x$ is the offset for the additive sharing of $x$ s.t. $x = \delta_x + \sum_i x^{\lir}$
		\end{itemize}

@@

		One challenge in this approach is the need to communicate the offset to the verifier so that they can verify the computations. However, a recent technique called \emph{Correlated GGM} (cGGM) has been introduced in \cite{GYWZ+22}, which allows the prover to generate random values $x^{\lir}$ using a modified GGM tree, such that $\sum_{i=1}^{N} x^{\lir} = x$, with no computational overload.
		This is achieved by enforcing an invariant over the cGGM tree: at each level, the sum of the nodes remains constant. More precisely, given a parent node $x$, the left child is defined as $F(x)$ and the right child as $x - F(x)$, where $F$ is a cryptographic psuedorandom function. As a result, only approximately half of the tree nodes require PRF evaluations, compared to all nodes in the classical GGM construction.
		This technique not only reduces the computational cost but also eliminates the need to transmit the offset to the verifier, thereby improving the overall communication efficiency.

@@

		\subsection{SBC Problem}
		The problem is the following: let $q$ be a prime power and two positive integers $k$, $n$.
		\newline
		Given two vectors $\vec{u}, \vec{v} \in (\mathbb{F}_{q^k})^n$, which are linearly independent over $\mathbb{F}_q$, find two non-colinear vectors $\vec{x}, \vec{y} \in (\mathbb{F}_q)^n$ such that $$(\vec{u} \cdot \vec{x})(\vec{v} \cdot \vec{y}) = (\vec{u} \cdot \vec{y})(\vec{v} \cdot \vec{x})$$
		where $(\cdot)$ is the dot product between 2 vectors.
		We denote an instance of the SBC problem given by two vectors $\vec{u}, \vec{v} \in (\mathbb{F}_{q^k})^n$ by $\text{SBC}[\vec{u}, \vec{v}]$. If the vectors $\vec{x},\vec{y} \in (\mathbb{F}_q)^n$ are solution of $\text{SBC}[\vec{u},\vec{v}]$, we use the notation $(\vec{x}, \vec{y}) \in \text{SBC}[\vec{u}, \vec{v}]$.

		\textbf{NSBC variant}.
		The authors also introduce a particular version of this problem called normalized SBC (NSBC) where $\vec{x}, \vec{y} \in (\mathbb{F}_q)^n$ and
		$\vec{x} = (\vec{x'}, 1, 0), \vec{y} = (\vec{y'}, 0, 1)$ to ensure that they are not colinear without needing to regenerate each time they are not. We also denote $(\vec{x}, \vec{y}) \in \text{NSBC}[\vec{u}, \vec{v}]$ if $(\vec{x}, \vec{y})$ are solutions to the NSBC problem on $(\vec{u}, \vec{v})$.

		\textbf{Key generation}. The private key $(\vec{x}, \vec{y})$ is constructed as such: uniformly at random choose $\vec{x'}, \vec{y'} \in (\mathbb{F}_q)^{n-2}$ and set $\vec{x} = (\vec{x'}, 1, 0)$ and $\vec{y} = (\vec{y'}, 0, 1)$.
		For the public key, randomly choose $\vec{u} \in (\mathbb{F}_{q^k})^{n}$ and $v_1,\dots,v_{n-1} \in \mathbb{F}_{q^k}$.
		Finally, set $v_n$ as:
		$$
			v_n = \frac{(\vec{u} \cdot \vec{y})(\sum_{i=1}^{n-1}v_ix_i) - (\vec{u} \cdot \vec{x})(\sum_{i=1}^{n-1}v_iy_i)}{y_n(\vec{u} \cdot \vec{x})}.
		$$

@@

		In our setting, we consider bilinear systems that are slightly overdetermined. We argue that when using hybrid methods such as \emph{Crossbred}, which avoid the need to construct Macaulay matrices up to high degrees—as required by algorithms like F5—the only reductions to zero that are likely to occur are those explained by the F5 and Frobenius criteria. Indeed, other types of reductions to zero typically appear at degrees $n_x + 1$ and $n_y + 1$ for bilinear systems, which are not reached in our practical use cases. Consequently, we consider these systems to behave in a semi-regular manner, although we believe they are, in practice, easier to solve.

>>>>

	\newpage
	\tableofcontents
	\newpage
	\section{MPC-in-the-Head and SBC signature scheme}\label{sec2}

@@

		The \emph{MPC-in-the-Head} (MPCitH) framework, introduced by \cite{IKOS07}, is a technique for constructing zero-knowledge proofs. It enables a prover $\mathcal{P}$ to convince a verifier that they possess a secret witness $w$ for a given public input $x$, without revealing any information about the witness itself.

		Let $\mathcal{L}$ be a language in NP (i.e. $x \in \mathcal{L} \Longleftrightarrow \exists w, \mathcal{R}(x, w) = 1$). $\mathcal{R}$ is a NP-relation. Given a public $x$, $\mathcal{P}$ wants to prove that he knows $w$ such that: $\mathcal{R}(x, w)$.

@@

		We introduce some standard notations in MPCitH protocols. 
		\begin{itemize}
			\item $[w]$, $w$ is a shared value.
			\item $w^{\lir}$ is the $i$-th share of $w$.
			\item $w^{[e]}$ is a value used in the $e$-th round of the protocol.
			\item $\delta_w$ is the offset for the additive sharing of $w$ s.t. $w = \delta_w + \sum_i w^{\lir}$
		\end{itemize}

		To represent the secret witness $w$, a linear secret sharing scheme is used among $N$ simulated parties. Each party $i$ receives a random share denoted by $w^{\lir}$, and the prover computes an additional correction term $\delta_w$.

		This ensures that the full witness can be correctly reconstructed from the shares and the correction term during the simulated execution.

@@

@@

		One challenge in this approach is the need to communicate the offset to the verifier so that they can verify the computations. However, a recent technique called \emph{Correlated GGM} (cGGM) has been introduced in \cite{GYWZ+22}, which allows the prover to generate random values $x^{\lir}$ using a modified GGM tree, such that $\sum_{i=1}^{N} x^{\lir} = x$, with no computational overload.
		This is achieved by enforcing an invariant over the cGGM tree: at each level, the sum of the nodes remains constant. More precisely, given a parent node $x$, the left child is defined as $F(x)$ and the right child as $x - F(x)$, where $F$ is a cryptographic PRF. As a result, only approximately half of the tree nodes require PRF evaluations, compared to all nodes in the classical GGM construction.
		This technique not only reduces the computational cost but also eliminates the need to transmit the offset to the verifier, thereby improving the overall communication efficiency.

@@

		\subsection{Subfield Bilinear Collision Problem}
		The problem is the following: let $q$ be a prime power and two positive integers $k$, $n$.
		\newline
		Given two vectors $\vec{u}, \vec{v} \in (\mathbb{F}_{q^k})^n$, which are linearly independent over $\mathbb{F}_q$, find two non-colinear vectors $\vec{x}, \vec{y} \in (\mathbb{F}_q)^n$ such that 
		$$(\vec{u} \cdot \vec{x})(\vec{v} \cdot \vec{y}) = (\vec{u} \cdot \vec{y})(\vec{v} \cdot \vec{x}),
		$$
		where $(\cdot)$ is the dot product between 2 vectors.

		We denote an instance of the SBC problem given by two vectors $\vec{u}, \vec{v} \in (\mathbb{F}_{q^k})^n$ by $\text{SBC}[\vec{u}, \vec{v}]$. If the vectors $\vec{x},\vec{y} \in (\mathbb{F}_q)^n$ are solution of $\text{SBC}[\vec{u},\vec{v}]$, we use the notation $(\vec{x}, \vec{y}) \in \text{SBC}[\vec{u}, \vec{v}]$.

		\textbf{Normalized Subfield Bilinear Collision variant}.
		The authors also introduce a particular version of this problem called normalized SBC (NSBC) where $\vec{x}, \vec{y} \in (\mathbb{F}_q)^n$ and
		$\vec{x} = (\vec{x'}, 1, 0), \vec{y} = (\vec{y'}, 0, 1)$ to ensure that they are not colinear without needing to regenerate each time they are not.

		We also denote $(\vec{x}, \vec{y}) \in \text{NSBC}[\vec{u}, \vec{v}]$ if $(\vec{x}, \vec{y})$ are solutions to the NSBC problem on $(\vec{u}, \vec{v})$.
		\textbf{Key generation}. The private key $(\vec{x}, \vec{y})$ is constructed as follows: uniformly at random choose $\vec{x'}, \vec{y'} \in (\mathbb{F}_q)^{n-2}$ and set $\vec{x} = (\vec{x'}, 1, 0)$ and $\vec{y} = (\vec{y'}, 0, 1)$.
		For the public key, randomly choose $\vec{u} \in (\mathbb{F}_{q^k})^{n}$ and $v_1,\dots,v_{n-1} \in \mathbb{F}_{q^k}$.
		Finally, set $v_n$ as:
		$$
			v_n = \frac{(\vec{u} \cdot \vec{y})(\sum_{i=1}^{n-1}v_ix_i) - (\vec{u} \cdot \vec{x})(\sum_{i=1}^{n-1}v_iy_i)}{y_n(\vec{u} \cdot \vec{x})}.
		$$

@@

		In our setting, we consider bilinear systems that are slightly overdetermined. We argue that when using hybrid methods such as \emph{Crossbred}, which avoid the need to construct Macaulay matrices up to high degrees—as required by algorithms like F5—the only reductions to zero that are likely to occur are those explained by the F5 and Frobenius criterion. Indeed, other types of reductions to zero typically appear at degrees $n_x + 1$ and $n_y + 1$ for bilinear systems, which are not reached in our practical use cases. Consequently, we consider these systems to behave in a semi-regular manner, although we believe they are, in practice, easier to solve.
